The Driving Forces of API Evolution

### 2015-06-02 thesis/paper.tex
Added SPE-prog definitions

Software evolution theory suggests that changes to software must obey one of eight laws \cite{lehman90sview}, known as Lehman's laws. The laws are presented in Table \ref{table:expl_laws}. The laws are said to only apply to E-type systems, which are programs that solve problems or address applications in the real world. Two other types of systems are also identified; S-type systems and P-type systems. S-type systems are programs which are exactly derivable from a specification. An example of such a program is one which finds the lowest common denominator for two integers. An example of a P-type system is a chess program. These types of programs can be derived from a specification, but the rules of the program have to be abstracted. In the case of a chess program, the possible moves have to be abstracted, since exactly specifying all the possible permutations would be virtually impossible. 
%SPE stuff in here.

>>>>

Software evolution theory suggests that software programs can be grouped into three different types of systems, namely S-programs, P-programs and E-programs \cite{lehman1980programs}. S-programs are programs which are exactly derivable from a specification. An example of such a program is one which finds the lowest common denominator for two integers. A key characteristic of an S-program is that if its functionality is changed, the end product will be a new product and not an evolved version of the previous program, since the mapping between specification and program is exact. P-programs are programs which are inherently more complex, and ``solve a problem which can be precisely formulated, but whose solution must inevitably reflect an approximation of the real world''. Furthermore, P-programs are programs which are centered around the perception of users, analysts and/or programmers. This perception is often commonly a mathematical model of the problem. An example of such a program is a program which predicts weather, since it is derived from a set of hydrodynamic equations. E-programs are described as ``inherently even more change prone [than S-programs and P-programs]. They are programs that mechanize a human or societal activity''. These factors can be explained by the difference between P-programs and a E-programs which is that E-programs ``change the very nature of the problem to be solved'' by ``becoming a part of the world it models''. Further, a characteristic of E-programs is that, during development, the specification must be developed by involving predictions of the consequences of the deployment of that system will have, since once the program is deployed ``questions of correctness, appropriateness, and satisfaction arise which inevitably leads to additional pressure for change". Lehman's eight laws of software evolution are said to apply to such E-programs. The laws are presented in Table \ref{table:expl_laws}.
